### 2019-06-13 2-Chapters/1-Chapter/chapter1.tex
Rien de neuf

@@

>>>>

\TODOL{Ecrire les 1.1 et 1.2 en quelques paragraphes}

@@

We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly not stationary''.
This specific context makes it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms (that is why we relax the hypothesis and only consider up-to $M \leq K$ players in Chapter~\ref{chapter:5}).
We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
%
We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.

As explained before in Chapter~\ref{chapter:1}, the future IoT networks will require to support more and more communicating devices.
In this section, we show that intelligent devices in unlicensed bands can use MAB learning algorithms to improve spectral resource exploitation.
%
We evaluate the performance of two classical MAB learning algorithms, \UCB{} and Thomson Sampling, to handle the decentralized decision-making of Spectrum Access, applied to IoT networks.
We mean by \emph{decentralized} that learning is performed on the device side, and is spread on the totality of the devices in a network.
We also evaluate the learning performance when the number of intelligent end-devices grows.
%
We show that using learning algorithms
firstly extends devices battery autonomy (by reducing the packet-loss-ratio or equivalently, by improving the successful-communication rate),
and secondly does help to fit more devices in such networks, even when all end-devices are intelligent and are dynamically switching from channel to channel.
In the studied settings,
Stochastic MAB learning
% provides an up to $16\%$ gain in terms of successful transmission probabilities, and
has near optimal performance even in non-stationary settings with a majority of intelligent devices.

% \TODOL{Based on the publication, ``Multi-Armed Bandit Learning in IoT Networks and non-stationary settings'', see https://hal.inria.fr/hal-01575419}

The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
and the network usage is more and more fluctuatating.
% and the stochastic hypothesis is more and more violated.
To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions. Static devices form an interfering traffic, which could be generated by devices using other standards as well.
Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel. So our first hypothesis is actually not constraining.
We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
This second approach is ideal, but not realistic the most common situation of decentralized co-located IoT networks, and it is just used here as a reference.
Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.
